Nix for Polyglot, Reproducible Data Science Workflows

### 2025-10-23 paper.qmd
fixes #3

By *linked and executable code and data*, Peng referred to compiled source code
and runnable scripts. In this paper, I interpret this notion more broadly as the
*computational environment*: the complete set of software required to execute an
analysis. Here too, a continuum exists. At the minimal end, authors might only
name the main software used—say, the [R]{.proglang} programming language. More
careful authors might also specify the version of [R]{.proglang}, or list the
additional packages and their versions. Rarely, however, do authors specify the
operating system on which the analysis was performed, even though differences in
operating systems can lead to divergent results when using the same code and
software versions, as shown by @neupane2019. It is even less common for authors
to provide step-by-step installation instructions for the required software
stack.

@@

This entire example can be executed by running the `run_polyglot_example.sh`
script available in the root of the paper's repository.
This script automates the full process:
1.  It first executes `gen-env.R` inside a temporary [Nix]{.pkg} shell to build
    the `default.nix` file.
2.  It then executes `gen-pipeline.R` inside the newly defined environment. This
  triggers [Nix]{.pkg} to run the entire polyglot pipeline in the correct order.
Upon completion, the script will have generated all intermediate artifacts and
the final `readme.html` report found in the `pipeline-output` folder containing
the visualization.

>>>>

By *linked and executable code and data*, Peng referred to compiled source code
and runnable scripts. In this paper, I interpret this notion more broadly as the
*computational environment*: the complete set of software required to execute an
analysis. Here too, a continuum exists. At the minimal end, authors might only
name the main software used—say, the [R]{.proglang} programming language. More
careful authors might also specify the version of [R]{.proglang}, or list the
additional packages and their versions. Rarely, however, do authors specify the
operating system on which the analysis was performed, even though differences in
operating systems can lead to divergent results when using the same code and
software versions, as shown by @neupane2019. It is even less common for authors
to provide step-by-step installation instructions for the required software
stack, an omission often driven by institutional constraints. Journals, for
instance, can inadvertently work against reproducibility by imposing strict page
or word limits that leave no room for the necessary technical documentation,
effectively discouraging thoroughness in the name of brevity.

@@

This entire example can be executed by running the `run_polyglot_example.sh`
script available in the root of the paper's repository.
This script automates the full process:
1.  It first executes `gen-env.R` inside a temporary [Nix]{.pkg} shell to build
    the `default.nix` file.
2.  It then executes `gen-pipeline.R` inside the newly defined environment. This
  triggers [Nix]{.pkg} to run the entire polyglot pipeline in the correct order.

Upon completion, the script will have generated all intermediate artifacts and
the final `readme.html` report found in the `pipeline-output` folder containing
the visualization.
